Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality

### 2015-06-26 content/contribution/CoDeR.tex
added figures for operators, fixed EBNF templating (moved EBNF environment definition to the template folder, moved include to the article template to avoid the pagebreak), added the Axiom environment missing from Nick's last commit

				This operator is semantically equivalent to the \emph{basic pattern match} of SPARQL applied to a stream, and to the \emph{alpha nodes} of the Rete pattern matching algorithm applied to semantic data.
				Like these operators, the triple-select is one-pass, and so is trivially interpreted as the unary stream-to-stream operator $\sigma_{(s,p,o)}$, where ${(s,p,o)}$ is the basic pattern to be matched.

@@

				The contents of the produced stream are characterised as those triple instances from the input stream whose triples match the basic pattern ${(s,p,o)}$.

@@

				This operator encompasses all of the one-pass built-in datatype functions of RIF-Core, and is also trivially interpreted as a unary stream-to-stream operator $\sigma_{f(a_1,\dots,a_n)}$, where $f$ is the datatype function to apply and each $a_i$, ${1 \leq i \leq n}$ is either a variable or a constant value to which $f$ is to be applied.

@@

@@

@@

@@

>>>>

				This operator is semantically equivalent to the \emph{basic pattern match} of SPARQL applied to a stream, and to the \emph{alpha nodes} of the Rete pattern matching algorithm \citep{forgy79} applied to semantic data.
				Like these operators, the triple-select is one-pass, and so is trivially interpreted as the unary stream-to-stream operator $\sigma_{triple}$, where ${triple}$ is the basic pattern to be matched.

@@

				The contents of the produced stream are characterised as those triple instances from the input stream whose triples match the specified basic pattern.
				\begin{figure}[t]
					\centering
					\includegraphics[width=0.45\textwidth]{basic-pattern-match}
					\caption{The triple-select operator.}
					\labelfig{continuous datalog: basic pattern match}
				\end{figure}

@@

				\begin{figure}[b]
					\centering
					\includegraphics[width=0.45\textwidth]{datatype-function}
					\caption{The operator encompassing all datatype functions of RIF-Core.}
					\labelfig{continuous datalog: datatype function}
				\end{figure}
				This operator encompasses all of the one-pass built-in datatype functions of RIF-Core, and is also trivially interpreted as a unary stream-to-stream operator $\sigma_{external}$, where ${external = f(a_1,\dots,a_n)}$, $f$ is the datatype function to apply and each $a_i$, ${1 \leq i \leq n}$ is either a variable or a constant value to which $f$ is to be applied.

@@

				\begin{figure}[t]
					\centering
					\includegraphics[width=0.45\textwidth]{graph-to-triple}
					\caption{The graph-to-triple operator.}
					\labelfig{continuous datalog: graph-to-triple}
				\end{figure}

@@

				\begin{figure}[b]
					\centering
					\includegraphics[width=0.45\textwidth]{instance-implication}
					\caption{The instance implication operator.}
					\labelfig{continuous datalog: instance implication}
				\end{figure}

@@

				This is similar to the CONSTRUCT clause of SPARQL, or the final step of \emph{conflict set resolution} within the Rete algorithm \citep{forgy79}.

@@

				\begin{figure}[t]
					\centering
					\subfloat[The SteM operator.]{\includegraphics[width=0.45\textwidth]{SteM}}\\
					\subfloat[A typical window-join.]{\includegraphics[width=0.45\textwidth]{window-join}}
					\caption{Exploded views of two-pass joining operators.}
					\labelfig{continuous datalog: SteM}
				\end{figure}

### 2015-06-26 content/contribution/R4.tex
added figures for operators, fixed EBNF templating (moved EBNF environment definition to the template folder, moved include to the article template to avoid the pagebreak), added the Axiom environment missing from Nick's last commit

\include{ebnf}
\begin{nestedsection}{R4: Rule-based Reasoning over RDF streams using Rete}{implementation}
	As a proof of concept, we have implemented R4: a Rule-based Reasoner for RDF-streams using Rete.
	Reasoning in R4 is performed continuously and incrementally as dataflow networks that work directly on RDF streams.
	The system architecture is shown in \reffig{R4-architecture}.
	\begin{figure}[t]
		\centering
		\includegraphics[width=0.45\textwidth]{R4-architecture}
		\caption{The system architecture of R4.}
		\labelfig{R4-architecture}
	\end{figure}

	A rule document containing any number of rules is submitted to the system.
	We have chosen RIF-Core to be the language in which rules can be represented, for being compatible with RDF and other semantic web standards.
	The rule document specifies any domain-specific rules as well as an entailment regime \emph{if any} for background reasoning.
	We also added a simple extension to RIF-Core to enable users to specify window constraints as part of their rules, as shown in Listing~\ref{lst-S-RIF-Core}.

\begin{figure*}
\begin{ebnf}
\Rule{Document}{
        \Optional{\NTerm{IRIMETA}} 
        \Token{Document} 
        \Token{(} 
        \Optional{\NTerm{Base}} 
        \ZeroOrMore{\NTerm{Prefix}} 
        \ZeroOrMore{\NTerm{Import}} 
        \Optional{\NTerm{Group}} 
        \Token{)}
}
\Rule{Base}{
        \Token{Base} 
        \Token{(} 
        \NTerm{ANGLEBRACKIRI} 
        \Token{)}
}
\Rule{Prefix}{
        \Token{Prefix} 
        \Token{(} 
        \NTerm{Name} 
        \NTerm{ANGLEBRACKIRI} \Token{)}
}
\Rule{Import}{
        \Optional{\NTerm{IRIMETA}} 
        \Token{Import} 
        \Token{(} 
        \NTerm{LOCATOR} 
        \Optional{\NTerm{PROFILE}} 
        \Token{)}
}
\Rule{Group}{
        \Optional{\NTerm{IRIMETA}} 
        \Token{Group} 
        \Token{(} 
        \ZeroOrMore{\OrList{\NTerm{RULE} \OrSep \NTerm{Group}}} 
        \Token{)}
}
\Rule{RULE}{
        \Optional{\NTerm{IRIMETA}}
        \Token{Forall}
        \OneOrMore{\NTerm{Variable}}
        \Token{(}
        \NTerm{CLAUSE}
        \Token{)}
}
\OrRule{
        \NTerm{CLAUSE}
}
\Rule{CLAUSE}{
        \OrList{\NTerm{Implies} \OrSep \NTerm{ATOMIC}}
}
\Rule{Implies}{
        \Optional{\NTerm{IRIMETA}}
        \OrList{\NTerm{ATOMIC} \OrSep \Token{And} \Token{(} \ZeroOrMore{ATOMIC} \Token{)}}
        \Token{:-}
        \NTerm{FORMULA}
}
\Rule{LOCATOR}{
        \NTerm{ANGLEBRACKURI}
}
\Rule{PROFILE}{
        \NTerm{ANGLEBRACKURI} 
        \Optional{\NTerm{Window}}
}
\Rule{Window}{
        \NTerm{Number}
        \NTerm{TimeUnit}
}
\Rule{TimeUnit}{
        \OrList{\Token{ms} \OrSep \Token{s} \OrSep \Token{m} \OrSep \Token{h}}
}
\end{ebnf}
\caption{Streaming RIF-Core Grammar}
\label{lst-S-RIF-Core}
\end{figure*}

@@

 		\begin{figure*}
 			\centering
 			\includegraphics[width=0.9\textwidth]{example-rete-network}
 			\caption{An example Rete network with corresponding (RDFS) rules.}
 			\labelfig{example-rete-network}
 		\end{figure*}
	\end{nestedsection}
\end{nestedsection}

>>>>

\begin{nestedsection}{R4: Rule-based Reasoning over RDF streams using Rete}{implementation}
	As a proof of concept, we have implemented R4: a Rule-based Reasoner for RDF-streams using Rete.
	Reasoning in R4 is performed continuously and incrementally as dataflow networks that work directly on RDF streams.
	The system architecture is shown in \reffig{R4-architecture}.
	\begin{figure}[b]
		\centering
		\includegraphics[width=0.45\textwidth]{R4-architecture}
		\caption{The system architecture of R4.}
		\labelfig{R4-architecture}
	\end{figure}

	A rule document containing any number of rules is submitted to the system.
	We have chosen RIF-Core to be the language in which rules can be represented, for being compatible with RDF and other semantic web standards.
	The rule document specifies any domain-specific rules as well as an entailment regime \emph{if any} for background reasoning.
	We also added a simple extension to RIF-Core to enable users to specify window constraints as part of their rules, as shown in \reffig{lst-S-RIF-Core}.

	\begin{figure*}[t]
		\centering
		\begin{ebnf}
			\Rule{Document}{
			        \Optional{\NTerm{IRIMETA}} 
			        \Token{Document} 
			        \Token{(} 
			        \Optional{\NTerm{Base}} 
			        \ZeroOrMore{\NTerm{Prefix}} 
			        \ZeroOrMore{\NTerm{Import}} 
			        \Optional{\NTerm{Group}} 
			        \Token{)}
			}
			\Rule{Base}{
			        \Token{Base} 
			        \Token{(} 
			        \NTerm{ANGLEBRACKIRI} 
			        \Token{)}
			}
			\Rule{Prefix}{
			        \Token{Prefix} 
			        \Token{(} 
			        \NTerm{Name} 
			        \NTerm{ANGLEBRACKIRI} \Token{)}
			}
			\Rule{Import}{
			        \Optional{\NTerm{IRIMETA}} 
			        \Token{Import} 
			        \Token{(} 
			        \NTerm{LOCATOR} 
			        \Optional{\NTerm{PROFILE}} 
			        \Token{)}
			}
			\Rule{Group}{
			        \Optional{\NTerm{IRIMETA}} 
			        \Token{Group} 
			        \Token{(} 
			        \ZeroOrMore{\OrList{\NTerm{RULE} \OrSep \NTerm{Group}}} 
			        \Token{)}
			}
			\Rule{RULE}{
			        \Optional{\NTerm{IRIMETA}}
			        \Token{Forall}
			        \OneOrMore{\NTerm{Variable}}
			        \Token{(}
			        \NTerm{CLAUSE}
			        \Token{)}
			}
			\OrRule{
			        \NTerm{CLAUSE}
			}
			\Rule{CLAUSE}{
			        \OrList{\NTerm{Implies} \OrSep \NTerm{ATOMIC}}
			}
			\Rule{Implies}{
			        \Optional{\NTerm{IRIMETA}}
			        \OrList{\NTerm{ATOMIC} \OrSep \Token{And} \Token{(} \ZeroOrMore{ATOMIC} \Token{)}}
			        \Token{:-}
			        \NTerm{FORMULA}
			}
			\Rule{LOCATOR}{
			        \NTerm{ANGLEBRACKURI}
			}
			\Rule{PROFILE}{
			        \NTerm{ANGLEBRACKURI} 
			        \Optional{\NTerm{Window}}
			}
			\Rule{Window}{
			        \NTerm{Number}
			        \NTerm{TimeUnit}
			}
			\Rule{TimeUnit}{
			        \OrList{\Token{ms} \OrSep \Token{s} \OrSep \Token{m} \OrSep \Token{h}}
			}
		\end{ebnf}
		\caption{Streaming RIF-Core Grammar}
		\labelfig{lst-S-RIF-Core}
	\end{figure*}

@@

 		\begin{figure*}[t]
 			\centering
 			\includegraphics[width=0.9\textwidth]{example-rete-network}
 			\caption{An example Rete network with corresponding (RDFS) rules.}
 			\labelfig{example-rete-network}
 		\end{figure*}
	\end{nestedsection}
\end{nestedsection}

### 2015-06-26 content/contribution/continuous-datalog.tex
added figures for operators, fixed EBNF templating (moved EBNF environment definition to the template folder, moved include to the article template to avoid the pagebreak), added the Axiom environment missing from Nick's last commit

\begin{axiom}
\label{axiom:continuous datalog: window range leq 0}
A window of range ${X \leq 0}$ over any stream $S$ at any instant ${i
  \in \mathbb{N}}$ is always empty.
\begin{equation*}
W^{S,X}_{i} = \varnothing
\end{equation*}
\end{axiom}

\begin{axiom}
\label{axiom:continuous datalog: window composition}
A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any
given instant ${i \in \mathbb{N}}$ is equivalent to the union of the
$X$ consecutive windows of range 1 over the stream $S$ preceding and
including that at instant $i$.
\begin{equation*}
W^{S,X}_{i} = \mathop{\cup}_{i=0}^{X-1} W^{S,1}_{i-i}
\end{equation*}
\end{axiom}

\begin{axiom}
\item\label{axiom:continuous datalog: window disjointness}
A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any
instant ${i \in \mathbb{N}}$ is disjoint from a second window of any
range ${Y \in \mathbb{N}}$ over $S$ at any instant after ${i - X}$.
\begin{equation*}
\forall X' \geq X \, \left( W^{S,X}_{i} \cap W^{S,Y}_{i-X'} = \varnothing \right)
\end{equation*}
\end{axiom}

@@

\begin{axiom}
\label{axiom:continuous datalog: entailment disjointness}
If two sets, $A_{inst}$ and $B_{inst}$, of axiom instances are
disjoint, then the any set $E_{inst}$ of axiom instances each
justified at least in part by $A_{inst}$ is necessarily disjoint from
any set $F_{inst}$ of axiom instances each justified at least in part
by $B_{inst}$, because entailed fact instances are distinct with
regards to the fact instances that justify them.
\begin{multline*}
\left( A_{inst} \cap B_{inst} = \varnothing \right) \wedge \left( A_{inst} \vDash E_{inst} \right) \\
\wedge \left( B_{inst} \vDash F_{inst} \right) \rightarrow \left( E_{inst} \cap F_{inst} = \varnothing \right)
\end{multline*}
\end{axiom}

@@

\begin{axiom}
\label{axiom:continuous datalog: positive window increment}
The window of range 1 at any given instant ${i \in \mathbb{N}}$ over
the stream $P$ entailed by a ${CDP}$ is defined as the set of facts
that are entailed by the state of the ${CDP}$ at instant $i$ that were
not entailed by the state of ${CDP}$ at the previous instant instance
${i-1}$.
			\begin{equation*}
				W^{P,1}_{i} = \left( A' \cup E_{i} \right) \setminus \left( A' \cup E_{i-1} \right) = E_{i} \setminus E_{i-1}
			\end{equation*}
\end{axiom}

\begin{axiom}
\label{axiom:continuous datalog: negative window increment}
			The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous instant instance ${i-1}$ that are no longer entailed by the state of ${CDP}$ at instant $i$.
			\begin{equation*}
				W^{N,1}_{i} = \left( A' \cup E_{i-1} \right) \setminus \left( A' \cup E_{i} \right) = E_{i-1} \setminus E_{i}
			\end{equation*}
\end{axiom}

\begin{axiom}
\label{axiom:continuous datalog: entailment precedes negation}
			Facts may only be negated by some part of their justification leaving the sliding window $W^{S,R}_{i}$, this necessarily being after they were entailed.
			As such, a window of any range ${X \in \mathbb{N}}$ over the stream of entailments $P$ at any given instant ${i \in \mathbb{N}}$ is disjoint from a window of any range ${Y \in \mathbb{N}}$ over the the stream of entailment negations $N$ at any instant after ${i - X}$
			\begin{equation*}
				\forall X' \geq X \left( W^{P,X}_{i} \cap W^{N,Y}_{i-X'} = \varnothing \right)
			\end{equation*}
\end{axiom}

>>>>

	\begin{axiom}\label{axiom:continuous datalog: window range leq 0}
		A window of range ${X \leq 0}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is always empty.
		\begin{equation*}
			W^{S,X}_{i} = \varnothing
		\end{equation*}
	\end{axiom}

	\begin{axiom}\label{axiom:continuous datalog: window composition}
		A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is equivalent to the union of the $X$ consecutive windows of range 1 over the stream $S$ preceding and including that at instant $i$.
		\begin{equation*}
			W^{S,X}_{i} = \mathop{\cup}_{i=0}^{X-1} W^{S,1}_{i-i}
		\end{equation*}
	\end{axiom}

	\begin{axiom}\label{axiom:continuous datalog: window disjointness}
		A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over $S$ at any instant after ${i - X}$.
		\begin{equation*}
			\forall X' \geq X \, \left( W^{S,X}_{i} \cap W^{S,Y}_{i-X'} = \varnothing \right)
		\end{equation*}
	\end{axiom}

@@

	\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
		If two sets, $A_{inst}$ and $B_{inst}$, of axiom instances are
		disjoint, then the any set $E_{inst}$ of axiom instances each
		justified at least in part by $A_{inst}$ is necessarily disjoint from
		any set $F_{inst}$ of axiom instances each justified at least in part
		by $B_{inst}$, because entailed fact instances are distinct with
		regards to the fact instances that justify them.
		\begin{multline*}
			\left( A_{inst} \cap B_{inst} = \varnothing \right) \wedge \left( A_{inst} \vDash E_{inst} \right) \\
				\wedge \left( B_{inst} \vDash F_{inst} \right) \rightarrow \left( E_{inst} \cap F_{inst} = \varnothing \right)
		\end{multline*}
	\end{axiom}

@@

	\begin{axiom}\label{axiom:continuous datalog: positive window increment}
		The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at instant $i$ that were not entailed by the state of ${CDP}$ at the previous instant instance ${i-1}$.
		\begin{equation*}
			W^{P,1}_{i} = \left( A' \cup E_{i} \right) \setminus \left( A' \cup E_{i-1} \right) = E_{i} \setminus E_{i-1}
		\end{equation*}
	\end{axiom}

	\begin{axiom}\label{axiom:continuous datalog: negative window increment}
		The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous instant instance ${i-1}$ that are no longer entailed by the state of ${CDP}$ at instant $i$.
		\begin{equation*}
			W^{N,1}_{i} = \left( A' \cup E_{i-1} \right) \setminus \left( A' \cup E_{i} \right) = E_{i-1} \setminus E_{i}
		\end{equation*}
	\end{axiom}

	\begin{axiom}\label{axiom:continuous datalog: entailment precedes negation}
		Facts may only be negated by some part of their justification leaving the sliding window $W^{S,R}_{i}$, this necessarily being after they were entailed.
		As such, a window of any range ${X \in \mathbb{N}}$ over the stream of entailments $P$ at any given instant ${i \in \mathbb{N}}$ is disjoint from a window of any range ${Y \in \mathbb{N}}$ over the the stream of entailment negations $N$ at any instant after ${i - X}$
		\begin{equation*}
			\forall X' \geq X \left( W^{P,X}_{i} \cap W^{N,Y}_{i-X'} = \varnothing \right)
		\end{equation*}
	\end{axiom}
